Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB

### 2022-02-04 doc/appendix-snow-sdmTMB.Rmd
finally get layout working for owl appendix

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 1,
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE
)
```

@@

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
library(sdmTMB)
library(sf)
```

@@

```{r mesh, echo=TRUE, cache=FALSE, out.width="70%"}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
plot(mesh)
```

@@

```{r qqnorm, out.width="33%"}
qqnorm(pred$resid)
qqline(pred$resid)
```

@@

```{r stan-qqnorm, out.width="33%"}
qqnorm(r_mcmc)
qqline(r_mcmc)
```

@@

```{r plot-all-effects, echo=TRUE, eval=TRUE, cache=FALSE}
ggplot(data = p_proj) +
  geom_point(data = p, aes(X, Y, colour = exp(est)), size = 0.4, alpha = 0.5) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    trans = "sqrt", # makes variation in low numbers visible
    # can also trim the extreme high values
    limits = c(0, quantile(exp(p$est), 0.995)), na.value = "yellow",
    guide = guide_colourbar(title.position = "top")
  ) +
  labs(colour = "Predicted counts") +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

```{r plot-spatial-effects, echo=TRUE, eval=TRUE}
ggplot(data = p_proj) +
  geom_point(data = p, aes(X, Y, colour = omega_s), alpha = 0.5) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top"
  )) +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
```

@@

```{r plot-spatiotemporal-effects, echo=TRUE, eval=TRUE}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = epsilon_st),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(title.position = "top")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

```{r plot-fixed-effects, echo=TRUE, eval=TRUE}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = exp(est_non_rf)),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    breaks = c(0.03, 0.05, 0.07),
    guide = guide_colourbar(title.position = "top")
  ) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

>>>>

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.asp = 1.5, 
  out.width="100%",
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE
)
```

@@

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(patchwork)
library(dplyr)
# remotes::install_github("pbs-assess/sdmTMB", dependencies = FALSE)
library(sdmTMB)
library(sf)
```

@@

```{r mesh, echo=TRUE, cache=FALSE, out.width="80%"}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
plot(mesh)
```

@@

```{r qqnorm, fig.asp = 1, out.width="33%", cache=FALSE}
qqnorm(pred$resid)
qqline(pred$resid)
```

@@

```{r stan-qqnorm, fig.asp = 1, out.width="33%", cache=FALSE}
qqnorm(r_mcmc)
qqline(r_mcmc)
```

@@

```{r plot-all-effects, echo=TRUE, eval=TRUE, cache=FALSE, cache=FALSE, out.width="95%"}
ggplot(data = p_proj) +
  geom_point(data = p, aes(X, Y, colour = exp(est)), size = 0.4, alpha = 0.5) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    trans = "sqrt", # makes variation in low numbers visible
    # can also trim the extreme high values
    limits = c(0, quantile(exp(p$est), 0.995)), na.value = "yellow",
    guide = guide_colourbar(title.position = "top")
  ) +
  labs(colour = "Predicted counts") +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

```{r plot-spatial-effects, echo=TRUE, eval=TRUE, cache=FALSE, out.width="95%"}
ggplot(data = p_proj) +
  geom_point(data = p, aes(X, Y, colour = omega_s), alpha = 0.5) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(
    direction = "horizontal", title.position = "top"
  )) +
  theme(legend.position = c(0.1, 0.1), axis.title = element_blank())
```

@@

```{r plot-spatiotemporal-effects, echo=TRUE, eval=TRUE, cache=FALSE, out.width="95%"}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = epsilon_st),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(guide = guide_colourbar(title.position = "top")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```

@@

```{r plot-fixed-effects, echo=TRUE, eval=TRUE, cache=FALSE, fig.asp = 1.5, out.width="100%"}
ggplot(data = filter(p_proj, year > 1978)) +
  geom_point(
    data = filter(p, year > 1978), aes(X, Y, colour = exp(est_non_rf)),
    size = 0.4, alpha = 0.5
  ) +
  coord_sf(
    xlim = c(min(p_proj$X) - 50000, max(p_proj$X) - 50000),
    ylim = c(min(p_proj$Y), max(p_proj$Y))
  ) +
  scale_colour_viridis_c(
    breaks = c(0.03, 0.05, 0.07),
    guide = guide_colourbar(title.position = "top")
  ) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  facet_wrap(~year_f)
```
